---
title: "Homework 2"
author: "Kairui Wang"
date: "2023-09-30"
output: github_document
---



```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

1. Clean the data of 'pols-month'

```{r, message = FALSE}
month_df = 
  tibble(
    month_num = 1:12,
    month_abb = month.abb,
    month = month.name
  )

pols_df = 
  read_csv("./data/pols-month.csv") |>
  separate(mon, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    president = recode(prez_gop, "0" = "dem", "1" = "gop", "2" = "gop")) |>
  left_join(x = _, y = month_df) |> 
  select(year, month, everything(), -day, -starts_with("prez")) 
```

2. Clean the data of 'snp'

```{r, message = FALSE}
snp_df =
    read_csv(
      "./data/snp.csv", 
      col_types = cols(date = col_date(format = "%m/%d/%y"))) |>
  separate(date, into = c("year", "month_num", "day"), convert = TRUE) |>
  mutate(
    year = if_else(year > 2023, year - 100, year)) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, close) 
```

3. Tidy the data of `unemployment`

```{r, message = FALSE}
unemployment_df = 
  read_csv("./data/unemployment.csv") |>
  rename(year = Year) |>
  pivot_longer(
    Jan:Dec, 
    names_to = "month_abb",
    values_to = "unemployment"
  ) |> 
  left_join(x = _, y = month_df) |> 
  select(year, month, unemployment)
```

4. Merge the three datasets

```{r merge_538, message = FALSE}
data_538 = 
  left_join(pols_df, snp_df) |>
  left_join(x = _, y = unemployment_df)

str(data_538)
```


## Problem 2

1. Creat 'homes_powered' variable 

```{r, message = FALSE}
mr_df =
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = 1, range = "A2:N586") |>
  janitor::clean_names() |>
  mutate(
    year = as.character(year),
    homes_powered = (weight_tons*500/30)
  )
```

2. Clean and organize the data for Professor Trash Wheel and Gwynnda

```{r, message = FALSE}
prof_df =
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = 2, range = "A2:M108") |>
  janitor::clean_names() |>
  mutate(
    year = as.character(year),
    homes_powered = (weight_tons*500/30)
  ) |> 
  mutate(type = 'professor trash wheel')

gwyn_df =
  read_excel("./data/202309 Trash Wheel Collection Data.xlsx", sheet = 4, range = "A2:L157") |>
  janitor::clean_names() |>
  mutate(
    year = as.character(year),
    homes_powered = (weight_tons*500/30)
  ) |> 
  mutate(type = 'gwynnda trash wheel')
```

```{r}
mr_df = mutate(mr_df, type = 'mr trash wheel')
```

3. Combine with the Mr. Trash Wheel dataset

```{r, message = FALSE}
trash_df = 
  full_join(mr_df, prof_df) |> 
  full_join(gwyn_df)
```

The number of observations in `mr_df` dataset is `r nrow(mr_df)`. The names of key variables are: `r names(mr_df)`.

The number of observations in `prof_df` dataset is `r nrow(prof_df)`. The names of key variables are: `r names(prof_df)`.

The number of observations in `gwyn_df` dataset is `r nrow(gwyn_df)`. The names of key variables are: `r names(gwyn_df)`.

The total weight of trash collected by Professor Trash Wheel is `r sum(pull(prof_df, weight_tons))` tons. The total number of cigarette butts collected by Gwynnda in July of 2021 is `r gwyn_df |> filter(year == 2021, month == 'July') |> pull(cigarette_butts) |> sum()`.



## Problem 3

1. Import, clean, and tidy the dataset of baseline demographics

```{r, message = FALSE}
baseline_df = 
  read_csv("./data_mci/MCI_baseline.csv", skip = 1 ) |>
  janitor::clean_names() |>
  mutate( 
    sex = as.factor(sex),
    apoe4 = as.factor(apoe4)
    ) |> 
  filter(age_at_onset > current_age | age_at_onset == '.')

```

The number of observations that had an age of onset for MCI was `r nrow(baseline_df)`. 

The average baseline age is `r baseline_df |> pull(current_age) |> mean() |> round(2)`. 

A proportion of `r ((sum(pull(baseline_df, sex) == 0 & pull(baseline_df, apoe4) ==1)/sum(pull(baseline_df, sex) == 0))*100) |> round(2)` % of women in the study are APOE4 carriers.


2.Import, clean, and tidy the dataset of longitudinally observed biomarker values

```{r, message = FALSE}
amyloid_df =
  read_csv("./data_mci/mci_amyloid.csv", skip = 1) |>
  janitor::clean_names() |>
  pivot_longer(
    time_2:time_8,
    names_to = 'time',
    values_to = 'amyloid_ratio'
  ) |>
  rename(id = study_id) |>
  mutate(amyloid_ratio = as.numeric(amyloid_ratio))
```

The number of observations is `r nrow(amyloid_df)`. A total of `r length(setdiff(pull(baseline_df, id), pull(amyloid_df, id)))` participants appear in only the baseline dataset, and a total of `r length(setdiff(pull(amyloid_df, id), pull(baseline_df, id)))` participants appear in only the amyloid dataset. It suggests that there are unique 'id' values in each dataset that are not shared between the two datasets.


3.Combine the demographic and biomarker datasets

```{r, message = FALSE}
MCI_df <- inner_join(baseline_df, amyloid_df, by = "id")
```

The resulting dataset has `r nrow(MCI_df)` observations and `r ncol(MCI_df)` variables.

The average age at baseline for this group was `r round(mean(pull(MCI_df, current_age)), digits=2)` years.

4.Saving the combined dataset as a csv file

```{r, message = FALSE}
write.csv(MCI_df, file = "data_mci/MCI_df.csv")
```
